Openness and computational reproducibility in plant pathology: where do we stand and a way forward

### 2022-07-04 analysis/paper/paper.Rmd
Correct title capitalisation
Edit paper for grammar and spelling

---
title: "Openness and Computational Reproducibility in Plant Pathology: Where do we Stand and a Way Forward"
author:
  - Adam H. Sparks:
      email: Adam.Sparks@dpird.wa.gov.au
      institute: DPIRD, USQ
      correspondence: true
  - Emerson M. Del Ponte:
      institute: UFV
  - Kaique S. Alves:
      institute: UFV
  - Zachary S. L. Foster:
      institute: OSU
  - Niklaus J. Grünwald:
      institute: ARS
institute:
  - DPIRD: Department of Primary Industries and Regional Development, Farming Systems Innovation, Perth WA 6000, Australia
  - USQ: University of Southern Queensland, Centre for Crop Health, Toowoomba Qld 4350, Australia
  - UFV: Departmento de Fitopatologia, Universidade Federal de Viçosa, Brazil
  - OSU: Department of Botany and Plant Pathology, Oregon State University, Corvallis OR 97331, USA
  - ARS: Horticultural Crops Research Unit, USDA Agricultural Research Service, Corvallis OR 97330, USA
output: 
  bookdown::word_document2:
     fig_caption: yes
     number_sections: FALSE
     reference_docx: "../templates/phytopath_template.docx"
     pandoc_args:
       - '--lua-filter=scholarly-metadata.lua'
       - '--lua-filter=author-info-blocks.lua'
       - '--lua-filter=pagebreak.lua'
bibliography: references.bib
link-citations: no
csl: "../templates/american-phytopathological-society.csl"
---

@@

Reproducibility is one component under the umbrella of open science.
By proactively practising open science, the work increases the chance to become more reproducible through the availability of data and code.
That is, open science leads to reproducibility and replicability.

@@

In an effort to understand where we stand as a discipline regarding open science and reproducible research, we surveyed a broad selection of articles to represent a wide swathe of publications to evaluate our collective status.
We hand-picked 21 journals which we felt represented research publications in the field of plant pathology (Table \@ref(tab:journals)) that encompassed a range of subject matter foci, applied and fundamental work, country of origin and ranking metrics, *e.g.* quartile range or citation index, to represent the discipline.
The aim was to gather as complete an overview of the status of computational reproducibility in plant pathology journals as possible and avoid bias in the findings by skewing towards high-impact journals that may have a greater influence.
From those 21 journals we randomly selected 450 articles published from 2012 to 2021.
Using R [@RCT2022], two lists were created, the first was a list of the 21 journals and the second was a list of the evaluators that were evaluating the articles for reproducibility.
Initially there were four evaluators, later a fifth was added and the list was recreated.
There were three evaluation events that took place to increase the yearly coverage of the evaluations.
During each scoring event, each evaluator was randomly assigned 50 unique articles to evaluate during an evaluation event.

@@

Depending on the scoring event, three to five of this paper's authors were each assigned to rate a randomised list of journal articles using scoring criteria devised for the purposes of this research in each of the three different scoring events as their time allowed for their participation.
In the first event, three authors (AS, EDP, and ZF) evaluated a total of 200 unique articles, in the second event two authors evaluated a total of 100 unique articles (EDP, KA) and in the third round, five authors (AS, EDP, KA, ZF, NG) evaluated a total of 150 unique articles for a total of 450 articles.
Each article was evaluated by only one evaluator save five articles randomly selected for inter-rater repeatability analysis from the third and final evaluation event.

@@

A custom function, `import_notes()`, was written to import the data and format the columns properly R [@RCT2022].
Values for software packages were checked for spelling consistency and corrections were made manually where necessary.
When working with these data in R, all strings of software character values were converted to fully upper-case to standardise the capitalisation and alleviate any issues with capitalisation used between evaluators.

@@

When predicting the availability of data availability as compared with *Phytopathology*, publications in two journals, *Phytopathologia Mediterrania* and *Molecular Plant Pathology*, can be considered clearly more likely to share data than publications in *Phytopathology*, but no effects were detectable in any other journal title (\@ref(fig:journal-parameters)B, Figure S2, Table S2).
The effect of *Molecular Plant Microbe Interactions* (Median = 0.67, 95% CI [-0.28, 1.57]) has a 92.15% probability of being positive (> 0), 90.61% of being significant (> 0.05), and 78.94% of being large (> 0.30).
The estimation successfully converged (Rhat = 1.000) and the indices are reliable (ESS = 25478).
The effect of *Phytopathologia Mediterranea* (Median = 1.25, 95% CI [0.32, 2.15]) has a 99.58% probability of being positive (> 0), 99.42% of being significant (> 0.05), and 97.73% of being large (> 0.30).
The estimation successfully converged (Rhat = 1.000) and the indices are reliable (ESS = 24986).

@@

Except for a few isolated cases, most papers were not fully computationally reproducible.
Very few authors choose to share both data and code.
More authors shared data due to journal requirements to share sequence data, but other types of data related to field experiments or other laboratory studies were not likely to be shared.
Code sharing was extremely rare, but in cases where it was shared, most were included as a part of the journal article's publication as extra materials rather than through data repositories.
The reasons for not sharing code or data were not clearly available from the papers themselves and so this work was unable to determine possible reasons for this situation.
We recognise that in some cases there may be commercial or intellectual property (IP) reasons for not sharing data or code, but these reasons should be clearly stated.
However, in most cases, the data are collected with public funding and the code was developed using similar funding and there are no commercial or IP issues that preclude sharing and in fact the funding agency may have guidelines in place for sharing these materials.

@@

There are different ways to make research more open and reproducible, which can be thought of as levels of reproducibility.
If a general workflow for producing academic research involves clearly defining a research question, obtaining data for testing the hypothesis, summarizing/analyzing and presenting data and results, writing the manuscript then the next logical steps are depositing the code used along with the raw data in proper repositories (Figure \@ref(fig:workflow-dia)).
Here we define three levels of reproducibility which are also related to the evolution of computational methods and reproducible practices and highlight papers that fit these definitions.

@@

Taking this approach even further, packaging the full analysis in a containerized software application, *e.g.*, Docker is a way to help ensure computational reproducibility but at the expense of added complexity.
Docker is an open-source containerization platform that enables users to package several applications and an operating system into containers, thereby standardising the executable components by combining application source code (or analyzes) with the operating system required to run that analysis on a user's computer.
However, there are drawbacks to using Docker.
It can be difficult to understand for a new user and new platforms, like the Apple M1 chip, may not be fully supported, which hinder the ability to share the container.
But, in most cases using an open-source language like R, Julia or Python allows us to share our work in a fashion where we know that the analysis will run exactly the same on every computer.
For more on using Docker for reproducible research, we would refer readers to @Nuest2020.
As an example of this approach, @Khaliq2020 provided a research compendium as a Docker container with a DOI and a full R package that enables readers to fully replicate their analysis of *Ascochyta rabiei* conidia dispersal in chickpea using the data collected and stepping through other points where weather data were investigated and various models were fit before deciding on the best fit and recreate any figures as published in the article [@Khaliq2020a].
When this level is employed, tools such as Binder, <https://mybinder.org>, can be used that allow readers and reviewers to launch an interactive session in their web browser and interact with the data and rerun the analysis in an RStudio instance or Jupyter notebooks as Mioroni et al. used [-@Miorini2018; -@Miorini2019].
@Kamvar2015 took a slightly different approach by including all files necessary for the analysis and most output files in a repository [@Kamvar2014] that also included an installable R package that was used for the original analysis.
While there are many other methods, these two approaches illustrate some of the best-practices where the data and other files were deposited in repositories with DOIs and reproducibility issues were addressed by using R packages to handle dependencies and other versioning issues making the work more portable.

@@

If steps are followed to make the data FAIR then it will be readable by humans and machines alike and this will help support discoveries and support further research.
In turn, sharing data will lead to new citations for your work as others discover and use it.
To make your data the most widely discoverable and usable, ensure that it has a persistent identifier.
A digital object identifier (DOI) is the most common (<https://www.doi.org/>) but the Handle.Net Registry (HNR) (<https://handle.net/>) is also an option.
There are different options for generating a DOI for your data and other materials.
FigShare, Zenodo and OSF.io all offer persistent archives along with a service to generate a DOI for your materials.
The use of a persistent identifier works to ensure that even if the data are moved, they can still be located using that unique identifier.
For more on FAIR data, visit Go-Fair <https://www.go-fair.org/fair-principles/>.

@@

distinct(rrpp, journal, IF_5year) %>%
  left_join(journal, by = "journal") %>% 
  rename(Journal = journal,
         `5 Year IF` = IF_5year) %>%
  flextable() %>%
  colformat_num(na_str = "NA") %>%
  autofit() %>%
  set_caption(
    "Journal titles selected for inclusion, their respective five year impact factors as of 2022 (5 Year IF) and the number (n) of articles from each journal that were evaluated."
  )
```

@@

```{r posteriors, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Posterior distribution visualizations for each of six models fitted to scoring data that were used to evaluate factors on reproducibility of 450 papers published in 21 plant pathology journals or plant pathology focused articles from other specialized journals. The journal title was tested for an effect on code availability (A) and data availability (B), the year of publication was tested for an effect on code availability (C) and data availability (D), and Five Year Impact Factor of Journal was tested for effect on code availability (E) and data availability (F). Model fitnesses were found to be good for all models.", cache=FALSE, fig.env='figure'}

@@

```{r journal-parameters, echo=FALSE, message=FALSE, warning=FALSE, fig.cap='95% credible interval for the probability associated with the journal title for (A) computational materials (code) and (B) data being made readily available to the public. A test for equivalence was unable to detect any clear differences between *Phytopathology* and all other journals sharing computational materials (code). However, a test for equivalence found that *Molecular Plant Pathology* and *Phytopathologia Mediterranea* were more likely have data available than articles found in *Phytopathology*. Intervals in grey have a median value less than that of *Phytopathology*, the base level used for the analysis and articles in these journals could be less likely to share computational materials and data than articles published in *Phytopathology*. Intervals in black have a median value that is greater than that of *Phytopathlogy*, the base level for the analysis and articles in these journals could be more likely to share data or code than those found in *Phytopathology*.', cache=FALSE, fig.env='figure'}

>>>>

---
title: "Openness and computational reproducibility in plant pathology: where do we stand and a way forward"
author:
  - Adam H. Sparks:
      email: Adam.Sparks@dpird.wa.gov.au
      institute: DPIRD, USQ
      correspondence: true
  - Emerson M. Del Ponte:
      institute: UFV
  - Kaique S. Alves:
      institute: UFV
  - Zachary S. L. Foster:
      institute: OSU
  - Niklaus J. Grünwald:
      institute: ARS
institute:
  - DPIRD: Department of Primary Industries and Regional Development, Farming Systems Innovation, Perth WA 6000, Australia
  - USQ: University of Southern Queensland, Centre for Crop Health, Toowoomba Qld 4350, Australia
  - UFV: Departmento de Fitopatologia, Universidade Federal de Viçosa, Brazil
  - OSU: Department of Botany and Plant Pathology, Oregon State University, Corvallis OR 97331, USA
  - ARS: Horticultural Crops Research Unit, USDA Agricultural Research Service, Corvallis OR 97330, USA
output: 
  bookdown::word_document2:
     fig_caption: yes
     number_sections: FALSE
     reference_docx: "../templates/phytopath_template.docx"
     pandoc_args:
       - '--lua-filter=scholarly-metadata.lua'
       - '--lua-filter=author-info-blocks.lua'
       - '--lua-filter=pagebreak.lua'
bibliography: references.bib
link-citations: no
csl: "../templates/american-phytopathological-society.csl"
---

@@

Reproducibility is one component under the umbrella of open science.
By proactively practicing open science, the work increases the chance to become more reproducible through the availability of data and code.
That is, open science leads to reproducibility and replicability.

@@

To understand where we stand as a discipline regarding open science and reproducible research, we surveyed a broad selection of articles to represent a wide swathe of publications to evaluate our collective status.
We hand-picked 21 journals which we felt represented research publications in the field of plant pathology (Table \@ref(tab:journals)) that encompassed a range of subject matter foci, applied and fundamental work, country of origin and ranking metrics, *e.g.*, quartile range or citation index, to represent the discipline.
The aim was to gather as complete an overview of the status of computational reproducibility in plant pathology journals as possible and avoid bias in the findings by skewing towards high-impact journals that may have a greater influence.
From those 21 journals we randomly selected 450 articles published from 2012 to 2021.
Using R [@RCT2022], two lists were created, the first was a list of the 21 journals and the second was a list of the evaluators that were evaluating the articles for reproducibility.
Initially there were four evaluators, later a fifth was added and the list was recreated.
There were three evaluation events that took place to increase the yearly coverage of the evaluations.
During each scoring event, each evaluator was randomly assigned 50 unique articles to evaluate during an evaluation event.

@@

Depending on the scoring event, three to five of this paper's authors were each assigned to rate a randomized list of journal articles using scoring criteria devised for the purposes of this research in each of the three different scoring events as their time allowed for their participation.
In the first event, three authors (AS, EDP, and ZF) evaluated a total of 200 unique articles, in the second event two authors evaluated a total of 100 unique articles (EDP, KA) and in the third round, five authors (AS, EDP, KA, ZF, NG) evaluated a total of 150 unique articles for a total of 450 articles.
Each article was evaluated by only one evaluator save five articles randomly selected for inter-rater repeatability analysis from the third and final evaluation event.

@@

A custom function, `import_notes()`, was written to import the data and format the columns properly R [@RCT2022].
Values for software packages were checked for spelling consistency and corrections were made manually where necessary.
When working with these data in R, all strings of software character values were converted to fully upper-case to standardize the capitalization and alleviate any issues with capitalization used between evaluators.

@@

When predicting the availability of data availability as compared with *Phytopathology*, publications in two journals, *Phytopathologia Mediterranea* and *Molecular Plant Pathology*, can be considered clearly more likely to share data than publications in *Phytopathology*, but no effects were detectable in any other journal title (\@ref(fig:journal-parameters)B, Figure S2, Table S2).
The effect of *Molecular Plant Microbe Interactions* (Median = 0.67, 95% CI [-0.28, 1.57]) has a 92.15% probability of being positive (> 0), 90.61% of being significant (> 0.05), and 78.94% of being large (> 0.30).
The estimation successfully converged (Rhat = 1.000) and the indices are reliable (ESS = 25478).
The effect of *Phytopathologia Mediterranea* (Median = 1.25, 95% CI [0.32, 2.15]) has a 99.58% probability of being positive (> 0), 99.42% of being significant (> 0.05), and 97.73% of being large (> 0.30).
The estimation successfully converged (Rhat = 1.000) and the indices are reliable (ESS = 24986).

@@

Except for a few isolated cases, most papers were not fully computationally reproducible.
Very few authors choose to share both data and code.
More authors shared data due to journal requirements to share sequence data, but other types of data related to field experiments or other laboratory studies were not likely to be shared.
Code sharing was extremely rare, but in cases where it was shared, most were included as a part of the journal article's publication as extra materials rather than through data repositories.
The reasons for not sharing code or data were not clearly available from the papers themselves and so this work was unable to determine possible reasons for this situation.
We recognize that in some cases there may be commercial or intellectual property (IP) reasons for not sharing data or code, but these reasons should be clearly stated.
However, in most cases, the data are collected with public funding and the code was developed using similar funding and there are no commercial or IP issues that preclude sharing and in fact the funding agency may have guidelines in place for sharing these materials.

@@

There are different ways to make research more open and reproducible, which can be thought of as levels of reproducibility.
If a general workflow for producing academic research involves clearly defining a research question, obtaining data for testing the hypothesis, analyzing, summarizing and presenting data and results, writing the manuscript then the next logical steps are depositing the code used along with the raw data in proper repositories (Figure \@ref(fig:workflow-dia)).
Here we define three levels of reproducibility which are also related to the evolution of computational methods and reproducible practices and highlight papers that fit these definitions.

@@

Taking this approach even further, packaging the full analysis in a containerized software application, *e.g.*, Docker is a way to help ensure computational reproducibility but at the expense of added complexity.
Docker is an open-source containerization platform that enables users to package several applications and an operating system into containers, thereby standardizing the executable components by combining application source code (or analyzes) with the operating system required to run that analysis on a user's computer.
However, there are drawbacks to using Docker.
It can be difficult to understand for a new user and new platforms, like the Apple M1 chip, may not be fully supported, which hinder the ability to share the container.
But, in most cases using an open-source language like R, Julia or Python allows us to share our work in a fashion where we know that the analysis will run exactly the same on every computer.
For more on using Docker for reproducible research, we would refer readers to @Nuest2020.
As an example of this approach, @Khaliq2020 provided a research compendium as a Docker container with a DOI and a full R package that enables readers to fully replicate their analysis of *Ascochyta rabiei* conidia dispersal in chickpea using the data collected and stepping through other points where weather data were investigated and various models were fit before deciding on the best fit and recreate any figures as published in the article [@Khaliq2020a].
When this level is employed, tools such as Binder, <https://mybinder.org>, can be used that allow readers and reviewers to launch an interactive session in their web browser and interact with the data and rerun the analysis in an RStudio instance or Jupyter notebooks as Mioroni et al. used [-@Miorini2018; -@Miorini2019].
@Kamvar2015 took a slightly different approach by including all files necessary for the analysis and most output files in a repository [@Kamvar2014] that also included an installable R package that was used for the original analysis.
While there are many other methods, these two approaches illustrate some of the best-practices where the data and other files were deposited in repositories with DOIs and reproducibility issues were addressed by using R packages to handle dependencies and other versioning issues making the work more portable.

@@

If steps are followed to make the data FAIR, then it will be readable by humans and machines alike and this will help support discoveries and support further research.
In turn, sharing data will lead to new citations for your work as others discover and use it.
To make your data the most widely discoverable and usable, ensure that it has a persistent identifier.
A digital object identifier (DOI) is the most common (<https://www.doi.org/>) but the Handle.Net Registry (HNR) (<https://handle.net/>) is also an option.
There are different options for generating a DOI for your data and other materials.
FigShare, Zenodo and OSF.io all offer persistent archives along with a service to generate a DOI for your materials.
The use of a persistent identifier works to ensure that even if the data are moved, they can still be located using that unique identifier.
For more on FAIR data, visit Go-Fair <https://www.go-fair.org/fair-principles/>.

@@

distinct(rrpp, journal, IF_5year) %>%
  left_join(journal, by = "journal") %>% 
  rename(Journal = journal,
         `5 Year IF` = IF_5year) %>%
  flextable() %>%
  colformat_num(na_str = "NA") %>%
  autofit() %>%
  set_caption(
    "Journal titles selected for inclusion, their respective five-year impact factors as of 2022 (5 Year IF) and the number (n) of articles from each journal that were evaluated."
  )
```

@@

```{r posteriors, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Posterior distribution visualizations for each of six models fitted to scoring data that were used to evaluate factors on reproducibility of 450 papers published in 21 plant pathology journals or plant pathology focused articles from other specialized journals. The journal title was tested for an effect on code availability (A) and data availability (B), the year of publication was tested for an effect on code availability (C) and data availability (D), and five-year impact factor of journal was tested for effect on code availability (E) and data availability (F). Model fitnesses were found to be good for all models.", cache=FALSE, fig.env='figure'}

@@

```{r journal-parameters, echo=FALSE, message=FALSE, warning=FALSE, fig.cap='95% credible interval for the probability associated with the journal title for (A) computational materials (code) and (B) data being made readily available to the public. A test for equivalence was unable to detect any clear differences between *Phytopathology* and all other journals sharing computational materials (code). However, a test for equivalence found that *Molecular Plant Pathology* and *Phytopathologia Mediterranea* were more likely have data available than articles found in *Phytopathology*. Intervals in grey have a median value less than that of *Phytopathology*, the base level used for the analysis and articles in these journals could be less likely to share computational materials and data than articles published in *Phytopathology*. Intervals in black have a median value that is greater than that of *Phytopathology*, the base level for the analysis and articles in these journals could be more likely to share data or code than those found in *Phytopathology*.', cache=FALSE, fig.env='figure'}
